\textbf{Socio-demographic determinants of COVID-19 vaccine uptake in Ontario: Exploring differences across the Health Region model}

### 2023-07-18 manuscript_v7/main.tex
new version from BN's comments

\date{}
\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, sharp corners, breakable]}{\end{tcolorbox}}\fi

@@

Covid-19, vaccination, survey, socio-economic factors, visible
minorities.

@@

Reasons given for this inequality in vaccination uptake have included
medical mistrust due to systemic medical
racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{17},\protect\hyperlink{ref-mosby2021}{22}},
mistrust in
vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{15}}, and
the existence of health misinformation and
disinformation\textsuperscript{\protect\hyperlink{ref-mosby2021}{22}--\protect\hyperlink{ref-freeman2020}{24}}.
However, it is important to consider that vaccination uptake can also
vary geographically across the administrative or political subdivisions
within a country. For example, differences in COVID-19 vaccination rates
have been associated with differences in attitudes towards vaccination
in administrative areas in New York and
Chicago\textsuperscript{\protect\hyperlink{ref-malik2020}{25}}, lack of
vaccine access in areas where underrepresented groups reside in
Ontario\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}, as well
as an absence of prioritization of areas inhabited by vulnerable groups
in the southeastern region of the
US\textsuperscript{\protect\hyperlink{ref-nguyen2021}{26}}. Other
studies have also shown heterogeneity in vaccine uptake within counties
in the
US\textsuperscript{\protect\hyperlink{ref-mollalo2021}{27}--\protect\hyperlink{ref-bhuiyan2022}{30}},
indicating that accounting for geographical differences in vaccination
can help predict patterns of booster uptake in Scottish
communities\textsuperscript{\protect\hyperlink{ref-wood2022}{31}}.
Therefore, the study of vaccination uptake across administrative of
political areas can be useful to highlight differences that governments
or public health agencies might need to address. However, in the case of
Canada, there is a limited amount of studies that have analyzed
geographical differences in vaccination. Existing studies in this area
have focused on differences within certain cities, such as
Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{32}}, or
Montréal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{33}}), or
have explored differences between
provinces-at-large\textsuperscript{\protect\hyperlink{ref-guay2022}{19}},
but to our knowledge, there are no studies that analyze differences in
vaccination uptake at the intra-provincial level. Such need is specially
important in the context of Canada's pandemic response goals, which have
been to minimize serious illness and deaths while minimizing societal
disruption\textsuperscript{\protect\hyperlink{ref-naci2023}{34}}.
Analyzing differences in vaccination uptake within the provinces can aid
to identify inequalities that might exist and that need to be addressed
before the advent of another pandemic.

@@

On the other hand, public health in Ontario is administered by Public
Health Ontario (PHO), a government agency established in late 2007 and
that is currently composed of 34 Public Health Units (PHUs) that cover
the entire geography of the province. Although PHUs were commissioned
with leading the distribution of the COVID-19 vaccine within their
respective
areas\textsuperscript{\protect\hyperlink{ref-rotenberg2021}{40}}, the
vaccine rollout occurred with significant interaction between PHO and
OH; in many instances both agencies had to work together to organize
vaccination clinics, with personnel associated with OH also being
actively involved when demand exceeded the capacity of PHO personnel,
especially in rural
areas\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}}. Indeed,
based on the experience of COVID-19, there is an ongoing debate on how
OH and PHO will interact in the future, and the challenges that this
will entail for the healthcare system of the
province\textsuperscript{\protect\hyperlink{ref-dion2022}{42}}.

@@

Therefore, in this study we hypothesized that there were differences in
COVID-19 vaccination rates between the Health Regions between October
2021 and January of 2022. To understand and contextualize these
differences, we by included socio-economic factors in our analysis,
aiming to identify in which demographic groups were particularly
impacted, in order to provide an assessment of the current state of
healthcare access in Ontario.

@@

We used data from the \emph{Survey of COVID-19 related Behaviours and
Attitudes}, a repeated cross sectional survey focused on the Canadian
province of Ontario that was commissioned by the Fields Institute for
Research in Mathematical Sciences and the Mathematical Modelling of
COVID-19 Task Force under ethical guidance from the University of
Toronto (under Research Ethics Board approval \#40999), and which ran
between September 30th, 2021 and January 17th,2022.

@@

Survey users entered their socio-economic information (age, income, and
racial/ethnic group), and asked information on vaccination status by
using the question ``Have you received the first dose of the COVID
vaccine?'', with possible answers ``yes'' and ``no''
(Table~\ref{tbl-descriptive-stats}). Of notice, because the
racial/ethnic categories of the survey did not match exactly the
categories from the 2016 Census, we used a combination of sources to
re-group certain categories in order to obtain estimates that could be
used to correct the data. Further details can be found in the Appendix.

Additionally, the survey automatically collected the geographical
location of the respondent (using the nearest municipality, as shown in
Figure~\ref{fig-map}), and the date of access to the survey. The
original dataset contained 39,029 observations. However, the number of
complete observations was much lower that the total number of
observations due to the survey design, which allowed respondents to exit
at any time and deployed the questions randomly. We selected the
observations with complete answers (6,343 observations, or 16.25\% of
the total) for our analysis. It should be noted that this response rate
is similar to response rates observed in previous studies that have used
the same type of survey instrument and that have ranged between 15\% and
22\%\textsuperscript{\protect\hyperlink{ref-kim2021}{45},\protect\hyperlink{ref-seeman2010}{46}}.
Next, we matched the city of each observation with its corresponding
LHIN and Health Region, and removed observations from areas with low
representation (254 observations corresponding to the North West and
North East Health Regions). After all the preliminary analyses, the
total number of observations used for analysis was 6,236 and included
the East, Central, Toronto, and West Health Regions covering between
October 1st, 2021 and January 17, 2022. The original dataset, clean
dataset, and details on the data cleaning process are described in
detail in the \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub
repository} for this paper.

@@

\hypertarget{statistical-analyses}{%
\subsection{Statistical analyses}\label{statistical-analyses}}

We used a logistic regression model to examine the impact of the Health
Regions in vaccination rates while considering the socio-economic
factors and months covered by the survey
(Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
Health Region and Race and income), as previous studies have shown that
socio-economic factors and their interactions are significant predictors
of intent of vaccination and vaccination
status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{47}--\protect\hyperlink{ref-cnat2022a}{49}}.
Because we identified differences in representativity between the survey
data and the estimates from the 2016 Census for the variables considered
in the analysis, and used used an iterative proportional fitting
procedure
(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{50}}
to correct the data using data from the Census and Health Region
population totals; and fitted the regression model to the uncorrected
and corrected data. Details regarding the correction can be found in the
Appendix. All analyses were conducted in R 4.2.2 using the packages
\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{51}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{52}},
\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{53}},
\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{54}},
and
\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{55}}.

@@

Figure~\ref{fig-models} presents the estimates (as odd ratios) from the
logistic regression models for vaccination status using the
socio-demographic factors collected by the survey, and their
interactions. Generally speaking, lower odds of vaccination were
identified in both cases in individuals characterized by a low household
income, or that identified as part of underrepresented groups. However,
the magnitude of the estimates differed between the uncorrected and
corrected models and more importantly, there were differences in the
statistical significance of certain estimates before and after the
correction. Specifically, the uncorrected model showed significant
differences in vaccination odds between the age groups considered, the
East Health Region, Latin American individuals with a household income
under CAD 25,000, and Indigenous individuals living in the Central
Health Region (Figure~\ref{fig-models},B) but these were deemed non
statistically significant after the correction.

@@

\caption{\label{fig-models}Coefficient estimates and confidence
intervals for the uncorrected model. Only statistically significant
interaction terms are shown. Full interaction terms can be found in
Supplementary Figures A-3 and A-4.}

@@

In this study, we hypothesized that there were differences in the uptake
of the first dose of the COVID-19 vaccine between individuals living in
the different Health Regions of Ontario during late 2021 and early 2022.
Using the Health Regions as the base of our analysis was advantageous,
as these newly implemented administrative areas match closely the
geographical boundaries of the Health Regions that have been
historically used to group Public Health Units in the
province\textsuperscript{\protect\hyperlink{ref-bolotin2018}{56},\protect\hyperlink{ref-ohesi}{57}}.
Thus, such analysis can provide information on vaccination rates with
relevance at both the public health and healthcare system level.
Moreover, such analysis could provide an overall assessment of
intra-provincial disparities that might need to be addressed moving
forward by decision-makers in order to ensure that in the event of a
future public health emergency or pandemic, OH and PHO will able to work
collaboratively in an efficient way having learned from the COVID-19
experience\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}},
thus ensuring that the population of the province benefits in the long
term from a robust health system.

@@

The implementation of the Health Regions in Ontario aimed at reducing
the bureaucratic complexity and health disparities identified under the
LHIN model. However, there are certain challenges that this new model
will need to overcome in order to work syngergistically with the rest of
the healthcare system of the province. First, each Health Region now
covers a large geographical area that was served in the past by multiple
LHINs. This in turn creates a complex socio-demographic landscape within
each Health Region that is likely to be different in each case due the
different levels of rurality and the proportion of underrepresented
groups in each Health Region.

Second, it is likely that a future pandemic or public health emergency
will impact the province in the years ahead. Although broadly speaking,
PHO and OH have been able to collaborate on the COVID-19 vaccination
rollout, future events will require a unified response from both the
public health arm of the government (represented by PHO), and the
healthcare system agency (OH). So far, the COVID-19 pandemic has served
as a stress test for both agencies, showing the necessity from decision
makers to clearly define how PHO and OH will interact in the future in
order to ensure that Ontarians can benefit by virtue of a robust
collaboration between both agencies. The evidence collected in this
study shows that healthcare disparities experienced by underrepresented
individuals in the province are the result of complex relationships
between infrastructure, resources, occupations, geography, and
socio-economic factors, among others. Therefore, this complexity
presents an ongoing challenge for the healthcare system of Ontario,
which will need to adapt and continue to aim for an integrative approach
that can serve those that are most vulnerable.

Finally, the recent nature in the adaption of the Health Region poses a
challenge for researchers in the acquisition of data and information
that can be used to analyze the performance of the new model. From one
side, the Health Regions have not been incorporated as part of Census
data (LHINs were considered before in the Census), and this impact the
amount and level of detail of available information. On the other hand,
evaluations of the Health Region model are at the moment not considered
in the Annual Reports of the Auditor General of Ontario, which was
critical in the past to identify limitations in the LHIN model.
Currently, the only demographic information available for each Health
Region is provided by Ontario Health (the agency that administers the
Health Regions) but this information only provides general estimates
that do not allow for detailed analyses of performance indicators (such
as hospitalizations, readmissions, and trends in chronic disease
incidence) between the Regions. Therefore, there is a pressing need for
open data that can be used by researchers and decision-makers to examine
the performance of the Health Region model.

The Health Region model will only by successful if it ensures that
healthcare improves across all segments of the population of Ontario,
particularly in the event of a future public health emergency or
pandemic where so far, based on the experience of the COVID-19 pandemic,
underrepresented individuals have been disproportionately affected.

>>>>

\date{}
\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

@@

Covid-19, vaccination, survey, socio-economic factors, visible
minorities, race and ethnicity

@@

Reasons given for this inequality in vaccination uptake have included
medical mistrust due to systemic medical
racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{17},\protect\hyperlink{ref-mosby2021}{22}},
mistrust in
vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{15}}, and
the existence of health misinformation and
disinformation\textsuperscript{\protect\hyperlink{ref-mosby2021}{22}--\protect\hyperlink{ref-freeman2020}{24}}.
However, it is important to consider that vaccination uptake can also
vary geographically across the administrative or political subdivisions
within a country. For example, differences in COVID-19 vaccination rates
have been associated with differences in attitudes towards vaccination
in administrative areas in New York and
Chicago\textsuperscript{\protect\hyperlink{ref-malik2020}{25}}, scarcity
of vaccination facilities in areas where underrepresented groups reside
in Toronto\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}, as
well as an absence of prioritization of areas inhabited by vulnerable
groups in the southeastern region of the
US\textsuperscript{\protect\hyperlink{ref-nguyen2021}{26}}. Other
studies have also shown heterogeneity in vaccine uptake within counties
in the
US\textsuperscript{\protect\hyperlink{ref-mollalo2021}{27}--\protect\hyperlink{ref-bhuiyan2022}{30}},
indicating that accounting for geographical differences in vaccination
can help predict patterns of booster uptake in Scottish
communities\textsuperscript{\protect\hyperlink{ref-wood2022}{31}}.
Therefore, the study of vaccination uptake across administrative of
political areas can be useful to highlight differences that governments
or public health agencies might need to address. However, in the case of
Canada, there is a limited amount of studies that have analyzed
geographical differences in vaccination. Existing studies in this area
have focused on differences within certain cities, such as
Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{32}}, or
Montréal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{33}}), or
have explored differences between
provinces-at-large\textsuperscript{\protect\hyperlink{ref-guay2022}{19}},
but to our knowledge, there are no studies that analyze differences in
vaccination uptake at the intra-provincial level. Such need is specially
important in the context of Canada's pandemic response goals, which have
been to minimize serious illness and deaths while minimizing societal
disruption\textsuperscript{\protect\hyperlink{ref-naci2023}{34}}.
Analyzing differences in vaccination uptake within the provinces can aid
to identify inequalities that might exist and that need to be addressed
before the advent of another pandemic.

@@

On the other hand, public health in Ontario is administered by Public
Health Ontario (PHO), a government agency established in late 2007 and
that is currently composed of 34 Public Health Units (PHUs) that cover
the entire geography of the province. Although PHUs were commissioned
with leading the distribution of the COVID-19 vaccine within their
respective
areas\textsuperscript{\protect\hyperlink{ref-rotenberg2021}{40}}, the
vaccine rollout occurred with significant interaction between PHO and
OH; in many instances both agencies had to work together to organize
vaccination clinics, with personnel associated with OH also being
actively involved when demand exceeded the capacity of PHO personnel,
especially in rural
areas\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}}. Indeed,
based on the experience of COVID-19, there are ongoing discussions on
how OH and PHO will interact in the future, and the challenges that this
will entail for the healthcare system of the
province\textsuperscript{\protect\hyperlink{ref-dion2022}{42}}.

@@

Therefore, in this study we wanted to understand if there were
differences in COVID-19 vaccination rates among the Health Regions
between October 2021 and January of 2022. To contextualize these
differences, we included socio-economic factors in our analysis aiming
to identify which demographic groups were particularly impacted, in
order to provide an assessment of the current state of healthcare access
in Ontario.

@@

We used data from the \emph{Survey of COVID-19 related Behaviours and
Attitudes}, a repeated cross-sectional survey focused on the Canadian
province of Ontario that was commissioned by the Fields Institute for
Research in Mathematical Sciences and the Mathematical Modelling of
COVID-19 Task Force under ethical guidance from the University of
Toronto (under Research Ethics Board approval \#40999), and which ran
between September 30th, 2021 and January 17th, 2022.

@@

Survey users entered their socio-economic information (age, income, and
racial/ethnic group), and were asked information on vaccination status
by using the question ``Have you received the first dose of the COVID
vaccine?'', with possible answers ``yes'' and ``no''
(Table~\ref{tbl-descriptive-stats}). Of notice, the racial/ethnic and
socio-economic categories from the survey did not match exactly the
categories from the 2016 Census (e.g., the categories ``East
Asian/Pacific Islander'' as well as the income bracket ``over CAD
110,000'' do not exist in the Census data). Therefore, we used a
combination of sources to re-group certain categories and socio-economic
strata in order to obtain estimates from the Census that could be used
to correct the data. Further details can be found in the Appendix.

Additionally, the survey automatically collected the geographical
location of the respondent (using the nearest municipality, as shown in
Figure~\ref{fig-map}), and the date of access to the survey. The
original dataset contained 39,029 observations. However, the number of
complete observations was much lower than the total number of
observations due to the survey design, which allowed respondents to exit
at any time and deployed the questions randomly. We selected the
observations with complete answers (6,343 observations, or 16.25\% of
the total) for our analysis. It should be noted that this response rate
is similar to response rates observed in previous studies that have used
the same type of survey instrument, with response rates between 15\% and
22\%\textsuperscript{\protect\hyperlink{ref-kim2021}{45},\protect\hyperlink{ref-seeman2010}{46}}.
Next, we matched the city of each observation with its corresponding
LHIN and Health Region, and removed observations from areas with low
representation (254 observations corresponding to the North West and
North East Health Regions). After all the preliminary analyses, the
total number of observations used for analysis was 6,236 and included
the East, Central, Toronto, and West Health Regions covering between
October 1st, 2021 and January 17, 2022. The original dataset, clean
dataset, and details on the data cleaning process, and data preparation
are described in detail in the
\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
for this paper.

@@

To analyze the survey data, we used a logistic regression model to
examine the impact of the Health Regions in the uptake of the first dose
of the vaccine, using as the variable of interest the question ``Have
you received the first dose of the COVID vaccine?'' (with possible
answers ``yes'' or ``no''), while considering as independent variables
the socio-economic factors of the participants, months covered by the
survey (Table~\ref{tbl-descriptive-stats}), and the Health Regions of
OH. Furthermore, we included certain interactions (Race and Health
Region and Race and income) in our model as previous studies have shown
that socio-economic factors and their interactions are significant
predictors of intent of vaccination and vaccination
status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{47}--\protect\hyperlink{ref-cnat2022a}{49}}.

Because we identified differences between the percentage of
participation within the socio-demographic factors of the survey
considered in in our analysis and the percentage within these factors in
the 2016 Census (such as participation within age groups and
racial/ethnic groups), we used an iterative proportional fitting
procedure
(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{50}}
to correct the percentages within each socio-demographic variable using
population totals obtained from the Census and OH; later, we fitted the
regression model to the uncorrected and corrected data to determine if
there were any differences in the obtanied estimates. Details regarding
the correction can be found in the Appendix. All analyses were conducted
in R 4.2.2 using the packages
\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{51}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{52}},
\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{53}},
\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{54}},
and
\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{55}}.

@@

Figure~\ref{fig-models} presents the estimates (as odd ratios) from the
logistic regression models for vaccination status using the
socio-demographic factors collected by the survey, and their
interactions for the corrected and uncorrected data. Generally speaking,
lower odds of vaccination were identified in both cases in individuals
characterized by a low household income, or that identified as part of
underrepresented groups. However, the magnitude of the estimates
differed between the uncorrected and corrected models and more
importantly, there were differences in the statistical significance of
certain estimates before and after the correction. Specifically, the
uncorrected model showed significant differences in vaccination odds
between the age groups considered, the East Health Region, Latin
American individuals with a household income under CAD 25,000, and
Indigenous individuals living in the Central Health Region
(Figure~\ref{fig-models},B) but these were deemed non statistically
significant after the correction.

@@

\caption{\label{fig-models}\textbf{A}. Coefficient estimates and
confidence intervals for the uncorrected model. \textbf{B}. Coefficient
estimates and confidence intervals for the corrected model. Only
statistically significant interaction terms are shown in both cases.
Full interaction terms can be found in Supplementary Figures A-3 and
A-4.}

@@

In this study, we wanted to understand if there were differences in
COVID-19 vaccination rates among the Health Regions between October 2021
and January of 2022 while identifying the socio-demographic determinants
(and their interactions) that could help predict these differences.
Using the Health Regions as the base of our analysis was advantageous,
as these newly implemented administrative areas match closely the
geographical boundaries of the Health Regions that have been
historically used to group Public Health Units in the
province\textsuperscript{\protect\hyperlink{ref-bolotin2018}{56},\protect\hyperlink{ref-ohesi}{57}}.
Thus, a Health Region-based analysis could provide information on
vaccination uptake that could be useful to understand patterns within
the realm of PHO and OH. Moreover, such analysis could provide an
overall assessment of intra-provincial disparities that might need to be
addressed moving forward by decision-makers in order to ensure that in
the event of a future public health emergency or pandemic, OH and PHO
will able to work collaboratively in an efficient way having learned
from the COVID-19
experience\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}},
thus ensuring that the population of the province benefits in the long
term from a robust health system.

@@

The evidence collected in this study shows that differences in the
uptake of the first dose of the COVID-19 vaccine by underrepresented
individuals in the province of Ontario were the result of complex
relationships between infrastructure, resources, occupations, geography,
and socio-economic factors. Specifically, we identified that between the
last trimester of 2021 and January of 2022 individuals that identified
as Arab or Middle Eastern, Black, Latin American, of a mixed background,
or that belonged to other racial groups had significantly lower odds of
vaccination when compared to White/Caucasian individuals in Ontario.

However, although overall vaccination uptake was lower in these
underrepresented individuals in the province as a whole, the odds of
vaccination differed within certain income levels and Health Regions in
these groups, specially for those that reported a low household income
and that had higher vaccination odds that individuals in the same
racial/ethnic group with a higher income. These results highlight the
complex landscape of the province, where varying degrees of rurality
exist in conjunction with a socio-demographic makeup that is unique to
each of the Health Regions.

Such complexity presents an ongoing challenge for the healthcare system
of Ontario, and in particular, to the Health Region model which will
need to adapt and build on the LHIN model with the aim of providing an
integrative approach that can serve those that are most vulnerable
across all the Health Regions of the province. In particular,
personalized care is an area that could be further developed to improve
vaccination uptake in underrepresented individuals in the future. As we
have shown above, there are currently a limited number of centers
focused on community-led healthcare in the province, and our results
showed the existence of differences in vaccination uptake between
certain Health Regions. Improving personalized care for underrepresented
individuals then, can serve as a point to improve trust and facilitate
vaccine access in these marginalized communities.

Furthermore, in the event of a future pandemic or public health
emergency in the province in the years ahead, PHO and OH will need to
collaborate on the implementation of public health measures (such as
vaccination campaigns) building on the experience gained during the
COVID-19 vaccination rollout. So far, the COVID-19 pandemic served as a
stress test for both agencies, showing the necessity from decision
makers to clearly define how PHO and OH will interact in the future in
order to ensure that Ontarians can benefit by virtue of a robust
collaboration between both agencies.

Finally, the recent nature in the adaption of the Health Region poses a
challenge for researchers in the acquisition of data and information
that can be used to analyze the performance of the new model. From one
side, the Health Regions have not been incorporated as part of Census
data (LHINs were considered before in the Census), and this impacts the
amount of data that can be obtained. On the other hand, evaluations of
the Health Region model are at the moment not considered in the Annual
Reports of the Auditor General of Ontario, which was critical in the
past to identify limitations in the LHIN model. Currently, the only
demographic information available for each Health Region is provided by
Ontario Health (the agency that administers the Health Regions) but this
information only provides general estimates that do not allow for
detailed analyses of performance indicators (such as hospitalizations,
readmissions, and trends in chronic disease incidence) between the
Regions. Therefore, there is a pressing need for open data that can be
used by researchers and decision-makers to assess the performance of the
Health Region model.
